
In Table \@ref(tab:sigTFs), the TFs with p-value <= 0.05 are shown.
`r num_sig_TFs` TFs are more often affected by the given SNPs than expected by the background sampling. 
All following tables display only information for the significantly identified TFs. 

```{r, echo = FALSE}

	#change R default string back to how it is in the result file
	sig_TFs$TF = gsub("..", "::", sig_TFs$TF, fixed=TRUE )
	sig_TFs$TF = gsub(".var.2.", "(var.2)", sig_TFs$TF, fixed=TRUE )
	sig_TFs$TF = gsub(".var.3.", "(var.3)", sig_TFs$TF, fixed=TRUE )

        knitr::kable(x = sig_TFs, format = "latex",longtable = T, booktabs = T, col.names = c( "TF", "observed TF hits", "# sampled rounds with higher or equal TF hits", "p-value"), align = "lccc", linesep = "", caption = "TFs significantly influenced by the given SNPs in comaprison to the random background sampling. The first column holds the TFs name, the second one how often a TFs binding sites was significantly changed by a SNP, the third one provides the number of sampled rounds with higher or equal TF hits than observed for the original data and the last column the resulting p-value.", row.names = F, label = "sigTFs") %>%
        kable_styling(latex_options = c("striped", "repeat_header")) %>%
	column_spec(1, width = "3cm")%>%
	column_spec(2, width = "2.5cm") %>%
	column_spec(3, width = "4cm") %>%
	column_spec(4, width = "2cm")
```

Table \@ref(tab:gainLossTFs) lists for all significantly identified TFs, if they are more often associated to an increasing (gain) or decreasing (loss) binding affinity comparing wild type (base first given in the input file) and the SNP risk allele. 
The p-value is computed using a two sided binomial test, assuming that gain and loss of a TF binding site occur equally likely (p = 0.5). We did not apply a p-value threshold on the table, however the predicted funtion (column 5) is only meaningful for a accordingly small p-values like 0.05. 
For instance, the first row of the table can be interpreted as following:
The binding sites of TF `r entries[1, 'TF']` are affected by in total `r as.double(entries[1, 'gain']) + as.double(entries[1, 'loss'])` SNPs.
Out of these `r entries[1, 'gain']` showed an increase in TF binding affinity (gain) and `r entries[1, 'loss']` a decrease (loss).
If binding sites of `r entries[1, 'TF']`  are affected it results more often in a `r entries[1, 'function']` of function (p-value `r entries[1, 'pvaluesBinomial']`).


```{r, echo = FALSE}

	#change R default string back to how it is in the result file
	entries$TF = gsub("..", "::", entries$TF, fixed=TRUE )
	entries$TF = gsub(".var.2.", "(var.2)", entries$TF, fixed=TRUE )
	entries$TF = gsub(".var.3.", "(var.3)", entries$TF, fixed=TRUE )

        knitr::kable(x = entries, format = "latex",longtable = T, booktabs = T, col.names = c( "TF", "gain", "loss", "p-value", "function"), align = "lcccc", linesep = "", caption = "Which TFs are associated to a gain or a loss of the TF binding affinity? The first column represents the TF name, the second and third ones hold the number of TF binding hits associated to a gain or loss, the fourth one shows the computed p-value and the last column the predicted function", digits = 4, row.names = F, label = "gainLossTFs") %>%
        kable_styling(latex_options = c("striped", "repeat_header")) %>%
	column_spec(1, width = "3cm")%>%
	column_spec(2, width = "1.5cm") %>%
	column_spec(3, width = "1.5cm") %>%
	column_spec(4, width = "2.5cm") %>%
	column_spec(5, width = "3cm")
```

